% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.

\addchap{Notation}
Below, we give an overview of the definitions and conventions commonly referred to in our paper.
\section*{Input}
\begin{itemize}
	\item $m\in\mathbb{N}$\ldots number of homogeneous servers
	\item $T\in\mathbb{N}$\ldots number of time slots
	\item $\lambda_1,\ldots,\lambda_{T}\in[0,m]$\ldots arrival rates
	\item $\Lambda\coloneqq(\lambda_1,\ldots,\lambda_T)$\ldots sequence of arrival rates
	\item $\beta\in\mathbb{R}_{\ge 0}$\ldots switching costs of a server
	\item $f:[0,1]\rightarrow\mathbb{R}$\ldots convex operating costs function of a server. Note that for Chapter~\ref{chap:approx_offline_scheduling}, $f$ is further assumed to be non-negative and monotonically increasing.	
	\item $\inp\coloneqq(m,T,\Lambda,\beta,f)$\ldots input of a problem instance
\end{itemize}

\section*{Problem statement}
\begin{itemize}
	\item $s_{i,t}\in\{0,1\}$\ldots state of server $i$ at time $t$, i.e.\ sleeping (0) or active(1)
	\item $S_i\coloneqq(s_{i,1},\ldots,s_{i,T})$\ldots sequence of states for server $i$
	\item $\lambda_{i,t}\in[0,1]$\ldots assigned load for server $i$ at time $t$
	\item $L_i\coloneqq(\lambda_{i,1},\ldots,\lambda_{i,T})$\ldots sequence of assigned loads for server $i$
	\item $\mathcal{S}\coloneqq(S_1,\ldots,S_m)$\ldots sequence of all state changes
	\item $\mathcal{L}\coloneqq(L_1,\ldots,L_m)$\ldots sequence of all assigned loads
	\item $\Sigma\coloneqq(\mathcal{S},\mathcal{L})$\ldots schedule for a problem instance $\inp$.
\end{itemize}

\section*{Miscellaneous}
\begin{itemize}
	\item $[n]$\ldots$\fromto{1}{n}$
	\item $x_t\in\fromto{0}{m}$\ldots number of active servers at time $t$
	\item $\mx\coloneqq(x_1,\ldots,x_T)$\ldots sequence of numbers of active servers. Note that due to Chapter~\ref{chap:preliminaries}, $\mx$ can be interpreted as a schedule.
	\item $\opcosts(x,\lambda)$\ldots costs of processing $\lambda$ with $x$ servers.
	\item $\swcosts(x_{t-1},x_{t})$\ldots costs of switching from $x_{t-1}$ to $x_t$ servers.
	\item $\costs(x_{t-1},x_{t},\lambda_t)$\ldots costs of switching from $x_{t-1}$ to $x_t$ servers and processing $\lambda_t$ with $x_t$ servers.
	\item $\costs(\mx)$\ldots costs of a schedule.
\end{itemize}

\section*{Conventions}
\begin{itemize}
	\item $\lambda_{t}=0$ for all $t\notin[T]$, i.e.\ there is no load before and after the scheduling process
	\item $s_{i,t}=0$ for all $t\notin[T]$, i.e.\ all servers are powered down before and after the scheduling process
\end{itemize}
